Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)

### 2023-05-08 doc/manuscript/ESSDmanuscript.tex

\begin{figure}
\includegraphics[width=14cm]{figures_tables/C_variable_mapping} \caption{\textbf{Schematic illustrating the carbon pools quantified under IPCC accounting; corresponding ForC variables, and relationships among them.} For each C pool, we show ForC variables corresponding to the stock, stock change (net annual increment), gain (influx), and loss. Most, but not all, EFDB-relevant ForC variables are shown here (excluded: C concentration in tree wood, annual branch turnover, understory biomass, total ecosystem C in all pools excluding soils and net annual increment thereof). Correspondence of ForC variables to IPCC criteria often depends upon measurement protocols (e.g., minimum stem diameter censused). Additional caveats are as follows: (a,b) branch fall and mortality of stems below the minimum stem diameter censused, which are necessary for a full accounting of dead organic matter production but typically assumed negligible for calculations of biomass change, are excluded by common measurement practice (a) or ForC variable definition (b); (c) assumes that leaf production equals leaf fall, or that changes in foliage biomass are negligble; (d,e) belowground components excluded by common measurement practice (d) or ForC variable definition (e); (f) excludes movement of dead wood into litter through breakage or size reduction; (g) measurements often limited to litter horizon (OL) and may exclude larger branches and stems classified as litter and/or the more decomposed layers of the O horizon.}\label{fig:fig_variable_mapping}
\end{figure}

@@

As of May 08, 2023, ForC (v4.0) contained 32693 independent records
(39855 total), 17293 of which were for the the 36 variables relevant to
EFDB (Fig. 1). These records were distributed across all forested
continents and ecozones, albeit unevenly (Fig. 2). The largest number of
records came from Asia, followed by North America, South America, and
Europe, with relatively few records from Africa, Australia, and Oceania
(Fig. 3c). Categorized by FAO ecozone, the greatest numbers of records
came from subtropical humid forests, temperate mountain systems, and
tropical rain forests, each with \textgreater2,000 independent records
(Fig. 3b). Boreal coniferous forests, temperate continental forests,
subtropical mountain systems, and tropical moist deciduous forests had
\textgreater1,000 independent records each, while other ecozones all had
\textless1,000 records. The most widely represented forest type was
needleleaf evergreen, followed by broadleaf deciduous and broadleaf
evergreen (Fig. 3a). In terms of stand age, the most represented age
class was 20-100 years, followed by \textless20 years and then
\textgreater100 years (Fig. 3d).

@@

ForC contained records for 22 of the 36 variables relevant to EFDB
(Table 2, Fig. 1). The records were very unevenly distributed across
variables. The variable with most records was aboveground biomass,
representing 45\% of all independent records relevant to EFDB, and
aboveground biomass components (woody biomass or foliage) representing
an additional 5\%. A total of 30\% of relevant records were for root
biomass (including fine and coarse root components), while 5\% described
total biomass. The non-living pools were less represented, with 5\% of
relevant were for dead wood (including standing and fallen components),
0.4\% for litter, and 2.3\% for soil carbon.

@@

There is vast potential to expand forest C data in EFDB by completing
the process of reviewing and sending data that are already in ForC
(Figs. 2-3). So far, only \textasciitilde8\% of the EFDB-relevant data
in ForC have been sent to EFDB.

>>>>

\begin{figure}
\includegraphics[width=14cm]{figures_tables/C_variable_mapping} \caption{\textbf{Schematic illustrating the carbon pools quantified under IPCC accounting; corresponding ForC variables, and relationships among them.} For each C pool, we show ForC variables corresponding to the stock, stock change (net annual increment), gain (influx), and loss. Most, but not all, EFDB-relevant ForC variables are shown here. Correspondence of ForC variables to IPCC criteria often depends upon measurement protocols (e.g., minimum stem diameter censused). Additional caveats are as follows: (a,b) branch fall and mortality of stems below the minimum stem diameter censused, which are necessary for a full accounting of dead organic matter production but typically assumed negligible for calculations of biomass change, are excluded by common measurement practice (a) or ForC variable definition (b); (c) assumes that leaf production equals leaf fall, or that changes in foliage biomass are negligble; (d,e) belowground components excluded by common measurement practice (d) or ForC variable definition (e); (f) excludes movement of dead wood into litter through breakage or size reduction; (g) measurements often limited to litter horizon (OL) and may exclude larger branches and stems classified as litter and/or the more decomposed layers of the O horizon.}\label{fig:fig_variable_mapping}
\end{figure}

@@

As of May 08, 2023, ForC (v4.0) contained 32693 independent records
(39855 total), 17293 of which were for the the 37 variables relevant to
EFDB (Fig. 1). These records were distributed across all forested
continents and ecozones, albeit unevenly (Fig. 2). The largest number of
records came from Asia, followed by North America, South America, and
Europe, with relatively few records from Africa, Australia, and Oceania
(Fig. 3c). Categorized by FAO ecozone, the greatest numbers of records
came from subtropical humid forests, temperate mountain systems, and
tropical rain forests, each with \textgreater2,000 independent records
(Fig. 3b). Boreal coniferous forests, temperate continental forests,
subtropical mountain systems, and tropical moist deciduous forests had
\textgreater1,000 independent records each, while other ecozones all had
\textless1,000 records. The most widely represented forest type was
needleleaf evergreen, followed by broadleaf deciduous and broadleaf
evergreen (Fig. 3a). In terms of stand age, the most represented age
class was 20-100 years, followed by \textless20 years and then
\textgreater100 years (Fig. 3d).

@@

ForC contained records for 23 of the 37 variables relevant to EFDB
(Table 2, Fig. 1). The records were very unevenly distributed across
variables. The variable with most records was aboveground biomass,
representing 45\% of all independent records relevant to EFDB, and
aboveground biomass components (woody biomass or foliage) representing
an additional 5\%. A total of 30\% of relevant records were for root
biomass (including fine and coarse root components), while 5\% described
total biomass. The non-living pools were less represented, with 5\% of
relevant were for dead wood (including standing and fallen components),
0.4\% for litter, and 2.3\% for soil carbon.

@@

There is vast potential to expand forest C data in EFDB by completing
the process of reviewing and sending data that are already in ForC
(Figs. 2-3). So far, only \textasciitilde8\% of the EFDB-relevant data
in ForC have been sent to EFDB. Moreover, the process of selecting
studies to review for transfer has yet to be optimized for
representation across geographic regions, forest types, and variables,
with the result that records transferred to date are unevenly
distributed across these categories (Figs. 2-3, Table 2). Future work
reviewing records for transfer will be most valuable if priority is
given to those from currently under-represented regions and forest
types, \emph{to records from countries relying on existing records (as
opposed to regular collection of new data through forest inventories)},
to the variables most needed by EFDB users, and to more contemporary
records.
